QuickSched: Task-based parallelism with dependencies and conflicts

### 2014-11-23 paper/paper.tex
added paragraph on Gadget-2.

\begin{figure}
    \centerline{\epsfig{file=figures/BH_scaling.pdf,width=0.9\textwidth}}
    \caption{Strong scaling and parallel efficiency of the Barnes-Hut tree-code
        computed over 1\,000\,000 particles.
        Solving the N-Body problem takes 323\,ms, achieving 75\% parallel
        efficiency, over all 64 cores.
        }
    \label{fig:BHResults}
\end{figure}

>>>>

For comparison, the same computations were run using the popular
astrophysics simulation software Gadget-2 \cite{ref:Springel2005},
using a traditional Barnes-Hut implementation based on octrees
and distributed-memory parallelism based on domain decompositions
and MPI \cite{ref:Snir1998}.
To achieve the same accuracy, an opening angle of 0.5 was used.
On a single core, the task-based tree traversal is already 1.9$\times$
faster than Gadget-2, due to the cache efficiency of the task-based
computations, which, by design, maximize the amount of computation
per memory access.
At 59 cores, where Gadget-2 performs best, the task-based tree traversal is
2.51$\times$ faster, and at the full 64 cores it is 4$\times$ faster,
due to the better strong scaling of the task-based approach as opposed
to the MPI-based parallelism in Gadget-2.

\begin{figure}
    \centerline{\epsfig{file=figures/BH_scaling.pdf,width=0.9\textwidth}}
    \caption{Strong scaling and parallel efficiency of the Barnes-Hut tree-code
        computed over 1\,000\,000 particles.
        Solving the N-Body problem takes 323\,ms, achieving 75\% parallel
        efficiency, over all 64 cores.
        For comparison, timings are shown for the same computation using
        the popular astrophysics code Gadget-2.
        The scaling for Gadget-2 (left) is shown relative to the performance of
        QuickSched, whereas the parallel efficiency (right) is computed relative
        to Gadget-2 on a single core.
        }
    \label{fig:BHResults}
\end{figure}
